### 2012-06-18 Weaklensing.tex
add Weaklensing.tex

%% LyX 2.0.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[utf8]{luainputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{babel}
\begin{document}

\part*{Weak Lensing Observables}

In weak lensing we can define complex ellipticity as $\epsilon=\frac{a-b}{a+b}e^{2i\phi}$
.a and b is long and short axis. While $\phi$is the orientation angle.

In analogy to the complex ellipticities we can also define complex
shear $\gamma=|\gamma|e^{2i\Phi}$as well as reduced shear $g=\frac{\gamma}{1-k}$.

The relation between unlensed and lensed ellipticity $\epsilon_{s}$
and $\epsilon$ is $\epsilon_{s}=\frac{\epsilon_{s}+g}{1+g*\epsilon}$.
In weak lensing limit,k$\ll1$and $\gamma\ll1$.Therefore $\epsilon\thickapprox\epsilon_{s}+\gamma$.

\part*{Method to obtain Shear $\gamma$}

Assume that we know the distribution of ellipticity P( \textbar{}$\epsilon_{s}|$)
in absence of lensing.

What we need to do is to find out the $\gamma$that leads to the \{$\epsilon_{s}\}$fits
P(\textbar{}$\epsilon_{s}|$) best.

We can use the $\chi^{2}$value to denote the deviation of \{$\epsilon_{s}$\}
from P(\textbar{}$\epsilon_{s}$\textbar{}).

Here $\chi^{2}=\underset{i}{\sum}\frac{(O_{i}-E_{i})^{2}}{E_{i}}$
$O_{i}$is the frequency of \{$|\epsilon_{s}|$\},while $E_{i}$is
the expected frequency obtained from P(\textbar{}$\epsilon_{s}|$).

Now that $\gamma-\chi^{2}$is a 2D function on complex plane. We can
use numerical methods like Simulated Anealing Method to calculate
the minimum point of this function.The $\gamma$ that leads to the
minimum $\chi^{2}$is the shear $\gamma$we want.
\end{document}