### 2012-06-19 K-S_test.tex
add K-S_test.tex

%% LyX 2.0.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{esint}
\usepackage{babel}
\begin{document}

\part*{Kolmogorov-Smirnov Test}

To test whether a set of data \{X\} satisfy probability distribution
P(X) ,we can use the K-S Test.

In K-S test $P=max(|F_{n}(x)-F(x)|)$ . $F_{n}(X)=($Observed values
NO. which is smaller than X in sample \{X\} )/N .N is the NO. of the
sample\{X\}.

F(X)=$\intop_{-\infty}^{x}P(X)$

The smaller P is ,the more likely \{X\} satisfies P(X).

In Weak Lensing for each $\gamma$ there's a set of ellipticity \{|$\epsilon_{s}|$\}.
We can use K-S Test to calculate the P value between \{|$\epsilon_{s}|$\}
and expected distribution P(|$\epsilon_{s}$|).

Then we can get a 2D $\gamma-P$ function. Find the minimum point
of this function, we obtain the shear $\gamma$.
\end{document}