### 2012-06-20 Likeilyhood.tex
add Likeilyhood.tex

%% LyX 2.0.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amsmath}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\DeclareRobustCommand{\greektext}{%
  \fontencoding{LGR}\selectfont\def\encodingdefault{LGR}}
\DeclareRobustCommand{\textgreek}[1]{\leavevmode{\greektext #1}}
\DeclareFontEncoding{LGR}{}{}
\DeclareTextSymbol{\~}{LGR}{126}

\makeatother

\usepackage{babel}
\begin{document}

\part*{Likeilyhood Function:}

Definition of likeilyhood function :

Let X be a random variable with a continuous probability distribution
with density function f depending on a parameter \textgreek{j}. Then
the function $L(\theta|x)=f_{\theta}(x)$ as a function of $\theta$is
the likeilyhood function of $\theta$,given the outcome x of X.

In a world ,the likelihood function of a set of parameter values given
some observed outcomes is equal to the probability of those observed
outcomes given those parameter values.

In the weak lensing case. We need to get the likeilyhood function
of shear $\gamma$,given a set of observed lensed ellipticity \{$\epsilon$\},and
we also know the distribution function of $|\epsilon_{s}$|,ellipticity
in absence of lensing P($|\epsilon_{s}|)$ .

Thus,the likeilyhood function of $\gamma$ should be $L(\gamma|\{\epsilon\})$=$\underset{i}{\prod}P($$|\epsilon_{s-n}|$).
Where $\epsilon_{s-n}=\epsilon_{-n}-\gamma$.

Or we can use the Log likeilyhood function $Log(L(\gamma|\{\epsilon\}))=\underset{i}{\sum}Log(P(|\epsilon_{s}|))$.
\end{document}