### 2013-03-11 docs/introduction.tex
Started MNRAS paper, with sections to be designed by Yike

%% LyX 2.0.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{textcomp}
\usepackage{babel}
\begin{document}

\part*{Introduction}

\section{Bias}

This bias on the shear is usually quantifi{}ed in terms of multiplicative
and additive errors This bias on the shear is usually quantifi{}ed
in terms of multiplicative and additive errors $m_{i}$ and $c_{i}$.
\begin{equation}
\hat{\gamma_{i}}=(1+m_{i})\gamma_{i}^{t}+c_{i}
\end{equation}

The requirements on the level of systematic errors for current and
future galaxy surveys are expressed in terms of mi, ci in Amara \&
R\textasciiacute{}efr\textasciiacute{}egier (2008).

For the simple average method,the systematic error of a shear measurement
method can be tested on a fi{}nite number of images in the absence
of shape noise by performing a \textquoteleft{}ring-test\textquoteright{}
(Nakajima \& Bernstein 2007).

For the traditional average method the bias of sear measurement comes
from the convolution with the PSF,pixelisation of the image , pixel
noise on the image due to the fi{}nite number of photons from the
source and atmosphere,and all the other effects that may cause error
in the galaxy ellipticity measurement.

(Tomasz Kacprzak\&Joe Zuntz 2012) present a method to effectively
remove this noise bias for realistic galaxy images. Using a forward
model fi{}tting, Maximum Likelihood approach for parameter estimation.
They create a model of the bias as a function of galaxy and PSF parameters
by determining their bias from various simulations that sample parameter
space. They demonstrate that this procedure successfully removes the
noise bias to the accuracy required by upcoming galaxy surveys.

\section{P(e) Intrinsic ellipticity distribution}

A number of studies have been made of the distribution of galaxy ellipticities
(e.g. ). These studies find a wide variation in distribution of |e|
, which appears strongly dependent on apparent magnitude.

(Leauthaud\&Massey 2007) undertake a measurement of the intrinsic
shape noise, $\sigma_{int}$, as a function of magnitude, size and
redshift,by derive an empirical estimation of the shape measurement
error, $\sigma_{measure}$.Then subtract it from $\sigma_{\gamma}^{2}$.
(The uncertanty of shear measurement in simple average is$\sigma_{\gamma}^{2}=\sigma_{int}^{2}+\sigma_{measure}^{2}$)

(http://arxiv.org/pdf/astro-ph/0702359v1.pdf)

(Joachimi\&Semboloni2012) study both the polarisation dispersion and
distribution of galaxy ellipticity using COSMOS data.

Their results shows a clear tendency towards lower $\sigma_{e}$ for
luminous galaxies with MV< \textminus{}22 in a broad redshift range
0.5 < z < 1.8.

They also plot P(|e|) for the full LRG sample and non-LRG galaxies
in the magnitude range \textminus{}19 < MV < \textminus{}17 ,based
on both COSMOS data and results from the simulation-based models.

\section{lensfit}

(Miller\& T. D. Kitchicung2007 ) proposed a modelfitting approach
in the galaxy shape measurement and evaluated how shear sensitivity
should be calculae based on posterior distribution of galaxies shape.

(Gamma estimaion:$\hat{g}=\frac{\sum w_{i}<e>_{i}}{\sum\partial<e>_{i}/\partial g}$
when the shape measurement is perfect thismeasurement goes back to
simple average)

It convolves the galaxy models with the PSF to calculate the posterior
probability of the model given the data then marginalize over galaxy
size, centroid and bulge fraction.The shear is then estimated in a
Bayesian way from this posterior probability surface.

To calculate the ellipticity prior which isrequired by the lensfit
method,(Miller\&T. D. Kitching2008) proposed an iterative mehod .(In
their model P(|e|) depends on 3 parameters and has a form cos{*}exp).They
proposed a model forP(|e|) and iteratively adjust three parameters
until the summed posterior and prior distributions agree.(? It looks
like they do not cares about the shear while calculating the ellipticity
posterior,then they use it in the shear estimator.?)

CFHTLenS is the fi{}rst weak lensing survey to apply the lensfi{}t
model fi{}tting method. 
\end{document}