### 2013-03-12 docs/formula.tex
add formula.tex

%% LyX 2.0.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{luainputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{esint}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{babel}

\makeatother

\usepackage{babel}
\begin{document}
Likelihood:

p($\epsilon_{\ell}|\vec{g},\alpha)=P(\epsilon_{0}(\epsilon_{\ell},\vec{g})|\alpha)*|\frac{\epsilon_{0}(\epsilon_{\ell},\vec{g})}{\epsilon_{\ell}}|$

Prior:

Prior($\vec{g})=1$ $|\vec{g}|<1$ else Prior(g)=0

Prior($\alpha)=exp(-\underset{i}{\xi\sum}(\alpha_{i}-\alpha_{i-1})^{2})$

Marginalized likelihood:

p($\epsilon_{\ell}|\vec{g})=\int p(\epsilon_{\ell}|\vec{g},\alpha)p(\alpha|\vec{g})d\alpha$

with $p(\alpha|\vec{g})=Prior(\alpha)$

$p(\epsilon_{l}|\vec{g})=\int p(\epsilon_{\ell}|\vec{g},\alpha)*Prior(\alpha)d\alpha$

$p(\epsilon_{\ell}|g_{1})=\underset{|g|<1}{\int}p(\epsilon_{\ell}|\vec{g})dg_{2}$
\end{document}